---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

# Introduction
This assignment is to analyze a person's daily activities in October and November in 2012. The data includes the number of steps taken in 5 minute intervals each day.

Of particular skills used - to convert character date variable to numeric date variable and plot the time series.  Minimum imputation is used to replace NAs with median.  Convert weekdays to 2 categories of workdays and offdays and to compare the activity pattern. Library(dplyr) is used for the most part for data management. Library(ggplot2) is used for visualization.

## Set Global Options
```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```

## Load Libraries
```{r load}

#remove.packages("rlang")
#install.packages("rlang")
#library(rlang)
#library(dplyr)

library(data.table)
library(lubridate)
library(dplyr)
library(ggplot2)

```

## Loading and preprocessing the data
```{r data}

act <- fread("activity.csv")

act$date <- as.Date(act$date, "%Y%m%d")
```

### Missing Value Count
There were a total of 2304 missing values.
```{r miss_cnt}

table(is.na(act$steps))

```

## What is mean total number of steps taken per day?
There were 61 days in October and November 2012.
```{r daily}

actdaily <- act %>% group_by(date) %>% 
  summarize(total_steps = sum(steps))

# str(actdaily)

```

## Summary of Total Steps by Date
```{r sum1}

summary(actdaily$total_steps)

## the mean is 10,766 steps
## the median is 10,765 steps
```

## What is the average daily activity pattern?
## Histogram of Daily Total Steps
```{r}

ggplot(actdaily, aes(x = total_steps)) +
    geom_histogram(fill = "blue", binwidth = 1000) +
    labs(title = "Daily Steps", x = "Daily Steps", y = "Frequency")

```

## Imputing missing values
## Activity Analysis with Imputation
```{r impute}

## impute NAs with median value
act$steps[is.na(act$steps)] <- median(act$steps, na.rm = T)

actdaily_imp <- act %>% group_by(date) %>% 
  summarize(total_steps = sum(steps))

```

## Are there differences in activity patterns between weekdays and weekends?
## Histogram of Daily Total Steps with Imputation
```{r imp1}

ggplot(actdaily_imp, aes(x = total_steps)) +
    geom_histogram(fill = "blue", binwidth = 1000) +
    labs(title = "Daily Steps", x = "Daily Steps", y = "Frequency")

```

## Summary of Total Steps with Imputation
```{r imp2}

summary(actdaily_imp$total_steps)

## Median is 10,395 steps

## Mean is 9,354 steps with imputation

```

## Comparison of Activity between Workdays and Offdays
```{r wday_comp}

act$date <- as.POSIXct(act$date, format = "%Y-%m-%d")

act$Weekdays <- weekdays(act$date)

#table(act$Weekdays)

act[grepl(pattern = "Monday|Tuesday|Wednesday|Thursday|Friday", x = Weekdays), "weekday_cat"] <- "workdays"

act[grepl(pattern = "Saturday|Sunday", x = Weekdays), "weekday_cat"] <- "offdays"
act[, `weekday_cat` := as.factor(`weekday_cat`)]

```

## Processed Data for Panel Plots
```{r pplot1}
##View(act)

analytic <- act %>% 
  group_by(interval, weekday_cat) %>% 
  summarize(avg_steps = mean(steps))

```

## Panel Plots for Workdays and Offdays
```{r pplot2}

ggplot(analytic, aes(x = interval, y = avg_steps, color = weekday_cat)) +
    geom_line() +
    facet_wrap(~ weekday_cat) +
    labs(title = "Workdays and Offdays Activity Comparison", 
             x = "Average Steps by Interval", 
             y = "Average Steps")

```

